Language and vision in conceptual processing: Multilevel analysis and statistical power

### 2022-06-29 manuscript/manuscript.Rmd
29062022

abstract: |
   Research over the past two decades has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the data: individual differences, words and tasks. To this aim, we drew on three existing data sets that were composed of large samples, and implemented the paradigms of semantic priming, semantic decision and lexical decision. After extending these data sets with language-based and vision-based measures, we performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. Overall, language-based information was found to be more important than vision-based information. Furthermore, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. In addition, higher-vocabulary participants presented a greater effect of language-based similarity than lower-vocabulary participants. Conversely, lower-vocabulary participants presented a greater effect of vision-based information, which was less relevant to the task. This 'relevance advantage' for higher-vocabulary participants also appeared in the semantic decision study, whose task required distinguishing between concrete and abstract words. Specifically, higher-vocabulary participants were more strongly influenced by word concreteness than lower-vocabulary participants. Furthermore, we compared the effects of different measures of visual information. Lastly, we estimated the sample size required to reliably examine each effect of interest. We found that 300 participants would suffice to examine language-based information contained in words, whereas more than 1,000 participants would be necessary to examine vision-based information as well as the interactions between both the former variables and vocabulary size, gender and presentation speed.

keywords: semantic, conceptual processing, vocabulary size, statistical power

@@

# Latex preamble specifying layout
header-includes: 
# Background on the first four lines: https://stackoverflow.com/a/70658799/7050882
  - \renewcommand\author[1]{}
  - \renewcommand\affiliation[1]{}
  - \authorsnames[1, 2, 2]{Pablo Bernabeu, Dermot Lynott, Louise Connell}
  - \authorsaffiliations{{Department of Psychology, Lancaster University, UK}, {Department of Psychology, Maynooth University, Ireland}}
  - \renewcommand{\topfraction}{.75}        # Reduce page margins
  - \renewcommand{\bottomfraction}{.4}      # Reduce page margins
  - \renewcommand{\textfraction}{.15}       # Reduce page margins
  - \renewcommand{\floatpagefraction}{.6}   # Reduce page margins
  - \setlength{\@fptop}{0pt}                # Vertically-align tables & figures to the top
  - \setcounter{topnumber}{3}               # Reduce page margins
  - \setcounter{bottomnumber}{2}            # Reduce page margins
  - \setcounter{totalnumber}{4}             # Reduce page margins
  - \usepackage{amsmath}                    # Maths formatting
  - \usepackage{upgreek}                    # Disable the italicisation of Greek letters
  - \usepackage{booktabs}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{longtable}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{array}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{multirow}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{wrapfig}                    # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{float}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{colortbl}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{pdflscape}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{tabu}                       # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttable}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttablex}            # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage[normalem]{ulem}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{makecell}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{xcolor}                     # Dependency for R package 'kableExtra' (table formatting)
  - \renewcommand\appendix{}                # Remove redundant label
---

@@

Research over the past two decades has suggested that conceptual processing involves both 'linguistic' and 'embodied' systems of the brain. The language system is characterised by associations across networks of words [@landauerIntroductionLatentSemantic1998; @pylyshynWhatMindEye1973], whereas the embodiment system is characterised by associations within perceptual, motor, affective and social domains [@barsalouPerceptualSymbolSystems1999; @diveicaQuantifyingSocialSemantics2022]. Both systems have strengths and limitations [@connell2019a; @dove2020a; @kumar2021a; @louwerse2015a; @mahon2016a]. As a result, conceptual processing flexibly draws on these systems following the demands of the context [@satoTaskRelatedModulation2008; @shebaniFlexibilityLanguageAction2018; @willemsFlexibilityEmbodiedLanguage2011].

The language system is involved in the representation of both abstract and concrete concepts [@ponari2018a; @regginDevelopmentAbstractWord2021]. For instance, measures of distributional semantic distance---solely based on texts---have been shown to capture aspects of the real world with considerable accuracy [@contreraskallensQuantifyingInterdisciplinarityCognitive2022; @dedeyneBetterExplanationsLexical2013; @kumarDistantConnectivityMultiplestep2020; @louwerseLanguageEncodesGeographical2009]. Although the language system is likely more important for the processing of abstract---rather than concrete---concepts [@barca2020a; @dunabeitiaQualitativeDifferencesRepresentation2009; @hultenNeuralRepresentationAbstract2021], although the degree of this difference could be small [@dedeyneBetterExplanationsLexical2013]. Language-based regularities may even allow distinguishing between concrete and abstract words, as word concreteness is consistently related to linguistic characteristics such as part of speech, morphological structure, countability and etymology [@lieversLinguisticDimensionsConcrete2021].

Regarding the time course, the language system is activated more quickly in semantically-deep tasks [@connell2013a; @louwerseTasteWordsLinguistic2011], whereas both systems are similarly fast in semantically-shallow tasks [@petilli2021a]. As a result of these characteristics, the language system is equipped to support a major part of conceptual processing [@connell2019a; @connellSeeHearWhat2014]. However, the language system is hindered by the symbol grounding problem, which refers to the impossibility of obtaining meaning---or 'grounding' lexical content in reality---on the sole basis of word-to-word associations [@guntherLimitsAutomaticSensorimotor2021; @guntherSymbolGroundingDirect2018; @louwerseSymbolInterdependencySymbolic2011; @harnadSymbolGroundingProblem1990]. Another limitation is caused by the finite vocabularies of languages, which sometimes fall short of the richness of sensorimotor experience [@guntherLimitsAutomaticSensorimotor2021; @dedeyneVisualAffectiveMultimodal2021; @majidOdorsAreExpressible2014; @majidSensesLanguageCulture2011; @majidCanNomenclatureBody2015]. In conclusion, word-to-word associations appear to be necessary but not sufficient for conceptual processing.

The embodiment system is best suited for the representation of concrete concepts, such as those involving perceptual features [@pecherDoesPizzaPrime1998; @yeeColorlessGreenIdeas2012; @ostarekTaskdependentCausalRole2017] and actions [@riccardiDissociatingActionAbstract2019]. Some research has suggested that embodied simulation plays an integral role in conceptual processing [@pulvermullerHowNeuronsMake2013], at least in the context of relatively-semantic tasks and stimuli [@leboisAreAutomaticConceptual2015; @pecherDoesPizzaPrime1998; @yeeColorlessGreenIdeas2012; @ostarekTaskdependentCausalRole2017; @connellWeKnowWhat2016; @garciaTimeFaceLanguage2020; @shebaniWhatImpairedSpatial2021; @vitaleNeuralInhibitionNetwork2021; @vukovicPrimaryMotorCortex2017; @wangSocialEmotionDimensional2021]. For instance, @riccardiDissociatingActionAbstract2019 investigated the comprehension of action verbs and abstract verbs in left-hemisphere stroke patients, and found that the processing of action concepts relied more heavily on sensory-motor areas. Furthermore, studies have suggested that even abstract concepts may be partially grounded in modality-specific domains such as emotion, sociality and interoception [@concaSearchDifferentCategories2021; @connellInteroceptionForgottenModality2018; @diveicaQuantifyingSocialSemantics2022; @kousta2011a; @vigliocco2014a; @wangSocialEmotionDimensional2021].

However, other research has curtailed the importance of this system [@mahonCriticalLookEmbodied2008; @bottiniConcretenessAdvantageLexical2021; @garceaPreservedToolKnowledge2013; @guntherSymbolGroundingDirect2018; @guntherLimitsAutomaticSensorimotor2021; @montero-melisNoEvidenceEmbodiment2022; @vannuscorpsPersistentSparingAction2016]. For instance, studies have found that lesions to motor brain areas did not systematically impair the processing of action concepts in language [@papeoActionPerformanceActionword2010; @negriWhatRoleMotor2007].

## Interplay between language and embodiment systems

The above characteristics make the language and the embodiment systems complementary to each other [@andrewsIntegratingExperientialDistributional2009; @dedeyneVisualAffectiveMultimodal2021; @wangNeuralRepresentationAbstract2010]. A comprehensive approach to both these systems requires a direct comparison of them. Studies tackling such a comparison have found that the systems are selectively engaged, following contextual demands [@connell2013a; @louwerseTasteWordsLinguistic2011; @ostarekTaskdependentCausalRole2017; @petilli2021a]. For instance, the role of perceptual simulation appears to be augmented in slower responses [@louwerseTasteWordsLinguistic2011], when words are presented more slowly [@lam2015a], and in tasks prompting deeper semantic processing [@connell2013a; @louwerseTasteWordsLinguistic2011; @ostarekTaskdependentCausalRole2017; @petilli2021a].

# The present study

There is considerable evidence supporting the 'hybrid' theory positing that conceptual processing depends on both language and embodiment systems [@barsalouLanguageSimulationConceptual2008; @louwerseSymbolInterdependencySymbolic2011; @connell2013a]. Nonetheless, there are also reasons to continue testing this theory. Firstly, the coexistence of several systems in a scientific theory must be thoroughly justified, due to the value of simplicity, as epitomised by Occam's razor [@galleseBrainConceptsRole2005; @tillmanHowSharpOccam2015]. This point is particularly pressing because the language system has consistently produced larger effect sizes than the embodiment system [@lam2015a; @pecherDoesPizzaPrime1998; @petilli2021a; @banksLinguisticDistributionalKnowledge2021; @kiela2014a; @louwerse2015a]. Secondly, mixed evidence has appeared regarding some aspects of the interplay. For instance, whereas some studies have suggested that the language system becomes relevant before the embodiment system [@lam2015a; @louwerseTasteWordsLinguistic2011], a recent study found the opposite pattern [@petilli2021a]. Similarly, whereas some evidence has suggested that female participants might draw on the language system more prominently than males [@hutchinson2013a], other research has suggested that this difference could be negligible [@wallentinChapterGenderDifferences2020]. Thirdly, the role of individual differences has not been examined in most of the studies addressing the language--embodiment interplay. Fourthly and last, the scarcity of statistical power that has ailed some studies in cognitive psychology and neuroscience [@vasishthHowEmbraceVariation2021; @marekReproducibleBrainwideAssociation2022] is likely to have affected some studies in the present topic area, too [see @lynottReplicationExperiencingPhysical2014; @montero-melisNoEvidenceEmbodiment2022].

The above open questions were addressed in the present study. These questions were overarched by the main research question, which examined whether the interplay between linguistic and visual processing held in the expected direction and at multiple levels of the experimental structure. To study this question, we drew on the data sets from three previous studies whose samples of participants were larger than average. These studies implemented the paradigms of semantic priming [@hutchison2013a], semantic decision [@pexman2017a] and lexical decision [@balota2007a]. Importantly, we extended these data sets with

@@

Experimental data in psycholinguistics often consists of various levels, such as participants, words and tasks. A simultaneous examination of these levels should improve our understanding of the evidence collected [@ostarekStrongInferenceResearch2021]. For instance, this cross-examination can help increase our certainty on the attribution of explanatory power to each level, as well as allowing a comparison of the explanatory power across levels. This multilevel approach is complementary to another approach in the field which seeks to test the causal contribution of different sources of information to conceptual processing---e.g., language [@ponari2018a], perception [@stasenkoWhenConceptsLose2014] and action [@speedImpairedComprehensionSpeed2017].

@@

## Participant level

*** to be edited *** Individual differences in domains such as language, perception, mental imagery and physical experience [e.g., @davies2017a; @dils2010a; @fettermanFeelingWarmBeing2018; @holt2006a; @mak2019a; @miceliDifferencesRelatedAging2022; @pexman2018a; @vukovic2015a; @yap2012a; @yap2017a].^[According to @lamiellStatisticalThinkingPsychology2019, 'individual differences' is a misnomer in that the methods used to analyse individual differences (e.g, regression) are not participant-specific. We think that this observation is mitigated in the context of mixed-effects models (as used in our current study), which incorporate by-participant random intercepts and random slopes.] Furthermore, in some topic areas in which individual differences have not featured as prominently, the data have revealed a non-trivial role of individual differences [@montero-melisConsistencyMotionEvent2021; @kosIndividualVariationLate2012].

@@

Vocabulary size refers to the number of words a person can recognise out of a discrete set. Higher-vocabulary participants are expected to respond faster overall [@pexman2018a]. Some previous studies have found that the effect of vocabulary size was moderated by variables related to general intelligence, such as processing speed [@ratcliff2010a; @yap2012a]. Owing to those findings, and the recommendations from other studies [@james2018a; @pexman2018a; @sassenhagenCommonMisapplicationStatistical2016], the present studies included covariates of vocabulary size, where available.These covariates were measures of intelligence that were not vocabulary-based, namely, attentional control in Study 1 and information uptake in Study 2 (such a variable was not available for Study 3).

@@

*** to be edited *** Semantic or lexical information in words [e.g., @dedeyneVisualAffectiveMultimodal2021; @lam2015a; @lund1995a; @lund1996a; @lynott2020a; @mandera2017a; @pexman2017a; @santosPropertyGenerationReflects2011; @wingfieldUnderstandingRoleLinguistic2022]

- *Variables in this study:* degree of association between words, degree of vision-based information in single words

- Covariates: lexical variables and word concreteness.

@@

Three hypotheses exist regarding the interaction between vocabulary size and lexical-semantic features. First, higher-vocabulary participants---compared to lower-vocabulary ones---might be more sensitive to linguistic features, thanks to a larger number of semantic associations [@connell2019a; @landauerIntroductionLatentSemantic1998; @louwerse2015a; @paivioMentalRepresentationsDual1990; @pylyshynWhatMindEye1973]. For instance, @yap2017a observed a larger semantic priming effect in higher-vocabulary participants from [@hutchinson2013a]. The second hypothesis, in contrast, states that higher-vocabulary participants would be *less* sensitive to linguistic features, thanks to a more automatic process. For instance, @yapIndividualDifferencesJoint2009 observed a smaller semantic priming effect in higher-vocabulary participants. Similarly, @yap2012a found that higher-vocabulary participants in a lexical decision task [@balota2007a] were less sensitive to a cluster of lexical and semantic features (i.e., word frequency, semantic neighborhood density and number of senses). Last, the third hypothesis proposes that higher-vocabulary participants might present a selective sensitivity, tailored to the task. For instance, @pexman2018a analysed the semantic decision task of @pexman2017a, in which participants assessed the concreteness of words. Pexman and Yap found that higher-vocabulary participants---compared to lower-vocabulary ones---were more sensitive to word concreteness and less sensitive to word frequency and age of acquisition, which were less relevant to the task [also see @lim2020a]. The three hypotheses herein were tested in the present study by revisiting some of the aforementioned studies [@hutchinson2013a; @pexman2017a; @balota2007a].

@@

Statistical power depends on the following factors: (1) sample size---comprising the number of participants, items, trials, etc.---, (2) effect size, (3) measurement variability and (4) number of comparisons being performed [@kumleEstimatingPowerGeneralized2021]. Out of these, sample size is the factor that can best be controlled by researchers [@kumleEstimatingPowerGeneralized2021]. The three studies we present below, containing larger-than-average sample sizes, offer an opportunity to conduct an a-priori power analysis to help determine the sample size of future studies [@albersWhenPowerAnalyses2018].

@@

## General method for statistical analyses {#general-statistical-analyses}

@@

Several covariates---or nuisance variables---were included, as detailed in each study, to allow a rigorous analysis of the effects of interest [@sassenhagenCommonMisapplicationStatistical2016]. Unlike the effects of interest, these covariates were not critical to the research question. They comprised lexical variables (e.g., word frequency) and participant-specific variables (e.g., attentional control). The lexical covariates were selected in every study out of the same 5 variables, which had been used as covariates in @wingfieldUnderstandingRoleLinguistic2022, and are widely used [e.g., @petilli2021a]. They comprised: number of letters (i.e., orthographic length), word frequency, number of syllables [both the latter from @balota2007a], orthographic Levenshtein distance [@yarkoniMovingColtheartNew2008] and phonological Levenshtein distance [@suarezObservingNeighborhoodEffects2011]. The selection among these candidates was performed because some of them were highly intercorrelated---i.e., $r$ > .70 [@dormannCollinearityReviewMethods2013; @harrison2018a]. The correlations and the selection models are available in [\underline{Appendix A}](#appendix-A-lexical-covariates).

@@

The participants and the stimulus items were crossed in the three studies. That is, each participant was presented with a subset of the stimulus words. Conversely, each word was presented to a subset of participants. Therefore, linear mixed-effects models were implemented. These models included a maximal random-effects structure, with by-participant and by-word random intercepts, and the appropriate random slopes for all effects of interest [@barrRandomEffectsStructure2013]. In the case of interactions, random slopes were included only when the variables interacting varied within the same unit [@brauer2018a]---e.g., an interaction of two variables varying within participants (only present in Study 1). Owing to convergence warnings, random slopes for controlled covariates were removed, inspired by Remedy 11 from @brauer2018a. Exceptionally, in the Bayesian models for the lexical decision study, the random slopes for the covariates did not have to be removed.

To avoid inflating the Type I error rate (false positives), the random slopes for the effects of interest were never removed [see Table 17 in @brauer2018a; for an example of this approach, see @diazNeuralSensitivityPhonological2021]. This approach provides a better protection against false positives [@barrRandomEffectsStructure2013; @brauer2018a; @singmann2019a] than the practice of removing random slopes when they do not significantly improve the fit [e.g., @bernabeu2017a; @pexman2018a; see @batesFittingLinearMixedeffects2015].

@@

*P* values were calculated with the Kenward-Roger approximation for degrees of freedom [@luke2017a], using the 'lmerTest' package, Version 3.1-3 [@R-lmerTest]. The latter package in turn used 'lme4', Version 1.1-26 [@batesPackageLme42021; @batesFittingLinearMixedeffects2015]. Diagnostics regarding convergence and normality are provided in [\underline{Appendix B}](#appendix-B-frequentist-analysis-diagnostics).

@@

In addition to the frequentist analysis, a Bayesian analysis was performed for two purposes: firstly, to enhance the interpretation of the smaller effects (which were identified as low-powered), and secondly, to complement the estimates obtained in the frequentist analysis. The latter purpose was pertinent because the frequentist models presented convergence warnings---even though it must be noted that a previous study found that frequentist and Bayesian estimates were similar despite convergence warnings appearing in the frequentist analysis [@rodriguez-ferreiroSemanticPrimingSchizotypal2020]. Furthermore, the complementary analysis was pertinent because the frequentist models presented residual errors that deviated from normality---even though mixed-effects models are fairly robust to such a deviation [@kniefViolatingNormalityAssumption2021; @schielzethRobustnessLinearMixed2020]. Owing to these precedents, we expected to find broadly similar estimates in the two approaches.

>>>>

abstract: |
   Research over the past two decades has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the data: individuals, words and tasks. To this aim, we drew on three existing data sets that were composed of large samples, and implemented the paradigms of semantic priming, semantic decision and lexical decision. After extending these data sets with language-based and vision-based measures, we performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. Overall, language-based information was found to be more important than vision-based information. Furthermore, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. In addition, higher-vocabulary participants presented a greater effect of language-based similarity than lower-vocabulary participants. In contrast, this pattern did not hold for vision-based information, which was less relevant to the task. The 'relevance advantage' in higher-vocabulary participants also appeared in the semantic decision and the lexical decision studies. Furthermore, we compare the effects of two visual information measures, and discuss the role of measurement instruments in this research topic. Lastly, we estimated the sample size required to reliably examine each effect of interest. We found that 300 participants suffice to examine language-based information contained in words, whereas more than 1,000 participants are necessary to examine vision-based information as well as the interactions between both the former variables and vocabulary size, gender and presentation speed.

keywords: conceptual processing, semantic priming, semantic decision, lexical decision, language, vision, vocabulary size, statistical power

@@

# Latex preamble specifying layout
header-includes: 
# Background on the first four lines: https://stackoverflow.com/a/70658799/7050882
  - \renewcommand\author[1]{}
  - \renewcommand\affiliation[1]{}
  - \authorsnames[1, 2, 2]{Pablo Bernabeu, Dermot Lynott, Louise Connell}
  - \authorsaffiliations{{Department of Psychology, Lancaster University, UK}, {Department of Psychology, Maynooth University, Ireland}}
  - \renewcommand{\topfraction}{.75}        # Reduce page margins
  - \renewcommand{\bottomfraction}{.4}      # Reduce page margins
  - \renewcommand{\textfraction}{.15}       # Reduce page margins
  - \renewcommand{\floatpagefraction}{.6}   # Reduce page margins
  - \setlength{\@fptop}{0pt}                # Vertically-align tables & figures to the top
  - \setcounter{topnumber}{3}               # Reduce page margins
  - \setcounter{bottomnumber}{2}            # Reduce page margins
  - \setcounter{totalnumber}{4}             # Reduce page margins
  - \usepackage{amsmath}                    # Maths formatting
  - \usepackage{upgreek}                    # Disable the italicisation of Greek letters
  - \usepackage{booktabs}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{longtable}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{array}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{multirow}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{wrapfig}                    # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{float}                      # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{colortbl}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{pdflscape}                  # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{tabu}                       # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttable}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{threeparttablex}            # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage[normalem]{ulem}             # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{makecell}                   # Dependency for R package 'kableExtra' (table formatting)
  - \usepackage{xcolor}                     # Dependency for R package 'kableExtra' (table formatting)
  - \renewcommand\appendix{}                # Remove redundant 'Appendix' label
---

@@

Over the past decades, research in psycholinguistics has suggested that conceptual processing depends on both language and embodiment systems [@barsalouLanguageSimulationConceptual2008; @louwerseSymbolInterdependencySymbolic2011; @connell2013a]. A comprehensive approach to both these systems requires a direct comparison of them. Studies tackling such a comparison have found that the systems are selectively engaged, following contextual demands [@connell2013a; @louwerseTasteWordsLinguistic2011; @ostarekTaskdependentCausalRole2017; @petilli2021a]. For instance, the role of perceptual simulation appears to be augmented in slower responses [@louwerseTasteWordsLinguistic2011], when words are presented more slowly [@lam2015a], and in tasks prompting deeper semantic processing [@connell2013a; @louwerseTasteWordsLinguistic2011; @ostarekTaskdependentCausalRole2017; @petilli2021a].

In spite of the amount of evidence favouring the interplay between language and embodiment, there are at least 2 reasons to continue testing the interplay theory. Firstly, the coexistence of several systems in a scientific theory must be thoroughly justified, due to the value of simplicity, as epitomised by Occam's razor [@galleseBrainConceptsRole2005; @tillmanHowSharpOccam2015]. This point is particularly pressing because the language system has consistently produced larger effect sizes than the embodiment system [@lam2015a; @pecherDoesPizzaPrime1998; @petilli2021a; @banksLinguisticDistributionalKnowledge2021; @kiela2014a; @louwerse2015a]. Consequently, in this research question, we examine whether the interplay between linguistic and visual processing holds in the expected directions and at multiple levels of the experimental structure.

Secondly, mixed evidence has appeared regarding some aspects of the interplay. For instance, whereas some studies have suggested that the language system is activated before the embodiment system [@lam2015a; @louwerseTasteWordsLinguistic2011], a recent study found the opposite pattern in a lexical decision task [@petilli2021a]. Similarly, some evidence has suggested that high-vocabulary participants are more sensitive to linguistic features [@yap2017a], whereas other evidence has suggested the opposite [@yapIndividualDifferencesJoint2009; @yap2012a]. Another case of mixed evidence regards gender: whereas some evidence has suggested that female participants draw on the language system more prominently than males [@hutchinson2013a], other research has suggested that this difference is negligible in the general population [@wallentinChapterGenderDifferences2020]. Presumably, the scarcity of statistical power that has ailed some studies in cognitive psychology and neuroscience [@vasishthHowEmbraceVariation2021; @marekReproducibleBrainwideAssociation2022] may have affected some studies in the present topic area as well [see @lynottReplicationExperiencingPhysical2014; @montero-melisNoEvidenceEmbodiment2022]. Therefore, in this study, we re-examine longstanding questions using larger-than-average data sets, and calculate the sample size required to reliably detect a range of effects.

## Language and vision

To retest the interplay theory, we revisit previous studies that contained larger-than-average samples of participants. Each of our three studies is organised around one *hub* study. In Study 1, the hub is @hutchison2013a, which implements a semantic priming paradigm. In Study 2, the hub is @pexman2017a, which implements a lexical decision paradigm. In Study 3, the hub is @balota2007a, which implements a lexical decision paradigm. Upon this basis, we added word-level variables from other studies to tap into language [@mandera2017a; @wingfieldUnderstandingRoleLinguistic2022] and vision [@lynott2020a; @petilli2021a]. Additionally, we included several covariates---or nuisance variables---to allow a rigorous analysis of the effects of interest [@sassenhagenCommonMisapplicationStatistical2016]. These covariates comprised participant-specific variables (e.g., attentional control), lexical variables (e.g., word frequency) and word concreteness.

@@

Experimental data in psycholinguistics often consists of various levels, such as participants, words and tasks. A simultaneous examination of these levels should enhance our understanding of the evidence collected [@ostarekStrongInferenceResearch2021]---for instance, by shedding light on the distribution of explanatory power within and across levels. This multilevel approach is complementary to another approach in the field that seeks to test the causal contribution of different sources of information to conceptual processing---e.g., language [@ponari2018a], perception [@stasenkoWhenConceptsLose2014] and action [@speedImpairedComprehensionSpeed2017].

@@

## Individual level

Variation  The role of individual differences in domains such as language, perception, mental imagery and physical experience [e.g., @davies2017a; @dils2010a; @fettermanFeelingWarmBeing2018; @holt2006a; @mak2019a; @miceliDifferencesRelatedAging2022; @pexman2018a; @vukovic2015a; @yap2012a; @yap2017a].^[According to @lamiellStatisticalThinkingPsychology2019, 'individual differences' is a misnomer in that the methods used to analyse individual differences (e.g, regression) are not participant-specific. We think that this observation is mitigated in the context of mixed-effects models (as used in our current study), which incorporate by-participant random intercepts and random slopes.] Furthermore, in some topic areas in which individual differences have not featured as prominently, the data have revealed a non-trivial role of individual differences [@montero-melisConsistencyMotionEvent2021; @kosIndividualVariationLate2012].

@@

Vocabulary size refers to the number of words a person can recognise out of a discrete set. Higher-vocabulary participants are expected to respond faster overall [@pexman2018a]. Some previous studies have found that the effect of vocabulary size was moderated by variables related to general intelligence, such as processing speed [@ratcliff2010a; @yap2012a]. Owing to those findings, and the recommendations from other studies [@james2018a; @pexman2018a; @sassenhagenCommonMisapplicationStatistical2016], the present studies included covariates of vocabulary size, where available. These covariates were measures of intelligence that were not vocabulary-based, namely, attentional control in Study 1 and information uptake in Study 2 (such a variable was not available for Study 3).

@@

*** to be edited *** Semantic or lexical information in words [e.g., @dedeyneVisualAffectiveMultimodal2021; @lam2015a; @lund1995a; @lund1996a; @lynott2020a; @mandera2017a; @petilli2021a; @pexman2017a; @santosPropertyGenerationReflects2011; @wingfieldUnderstandingRoleLinguistic2022]

- Variables in this study: language-based and vision-based information;

- Covariates: lexical variables and word concreteness;

@@

Three hypotheses exist regarding the interaction between vocabulary size and lexical-semantic features. First, higher-vocabulary participants---compared to lower-vocabulary ones---might be more sensitive to linguistic features, thanks to a larger number of semantic associations [@connell2019a; @landauerIntroductionLatentSemantic1998; @louwerse2015a; @paivioMentalRepresentationsDual1990; @pylyshynWhatMindEye1973]. For instance, @yap2017a observed a larger semantic priming effect in higher-vocabulary participants from [@hutchinson2013a]. The second hypothesis, in contrast, states that higher-vocabulary participants would be *less* sensitive to linguistic features, thanks to a more automatic process [@perfettiLexicalQualityHypothesis2002]. For instance, @yapIndividualDifferencesJoint2009 observed a smaller semantic priming effect in higher-vocabulary participants. Similarly, @yap2012a found that higher-vocabulary participants in a lexical decision task [@balota2007a] were less sensitive to a cluster of lexical and semantic features (i.e., word frequency, semantic neighborhood density and number of senses). Last, the third hypothesis proposes that higher-vocabulary participants might present a selective sensitivity, tailored to the task. For instance, @pexman2018a analysed the semantic decision task of @pexman2017a, in which participants assessed the concreteness of words. Pexman and Yap found that higher-vocabulary participants---compared to lower-vocabulary ones---were more sensitive to word concreteness and less sensitive to word frequency and age of acquisition, which were less relevant to the task [also see @lim2020a]. The three hypotheses herein were tested in the present study by revisiting some of the aforementioned studies [@hutchinson2013a; @pexman2017a; @balota2007a].

@@

Statistical power depends on the following factors: (1) sample size---comprising the number of participants, items, trials, etc.---, (2) effect size, (3) measurement variability and (4) number of comparisons being performed. Out of these, sample size is the factor that can best be controlled by researchers [@kumleEstimatingPowerGeneralized2021]. The three studies we present below, containing larger-than-average sample sizes, offer an opportunity to conduct an a-priori power analysis to help determine the sample size of future studies [@albersWhenPowerAnalyses2018].

@@

## General method for statistical analysis {#general-statistical-analyses}

@@

Several covariates---or nuisance variables---were included in each study, to allow a rigorous analysis of the effects of interest [@sassenhagenCommonMisapplicationStatistical2016]. Unlike the effects of interest, these covariates were not critical to the research question. They comprised participant-specific variables (e.g., attentional control), lexical variables (e.g., word frequency) and word concreteness.

The lexical covariates were selected in every study out of the same 5 variables, which had been used as covariates in @wingfieldUnderstandingRoleLinguistic2022 [also see @petilli2021a]. They comprised: number of letters (i.e., orthographic length), word frequency, number of syllables [both the latter from @balota2007a], orthographic Levenshtein distance [@yarkoniMovingColtheartNew2008] and phonological Levenshtein distance [@suarezObservingNeighborhoodEffects2011]. The selection among these candidates was performed because some of them were highly intercorrelated---i.e., $r$ > .70 [@dormannCollinearityReviewMethods2013; @harrison2018a]. The correlations and the selection models are available in [\underline{Appendix A}](#appendix-A-lexical-covariates).

Word concreteness was included due to its correlation with visual strength, as shown in each study below. A recent study suggested that the psycholinguistic import of concreteness is fundamentally lexical, and does not involve perceptual simulation [@bottiniConcretenessAdvantageLexical2021].

The participant-specific covariates were measures akin to fluid intelligence, and were included due to their relationship with vocabulary size, as previous studies have expressed the desirability of including such covariates [@james2018a; @pexman2018a]. As described in each study, these covariates were used in Studies 1 and 2, but not in Study 3, as such a variable was not available.

@@

The participants and the stimulus items were crossed in the three studies. That is, each participant was presented with a subset of the stimulus words. Conversely, each word was presented to a subset of participants. Therefore, linear mixed-effects models were implemented. These models included a maximal random-effects structure, with by-participant and by-item random intercepts, and the appropriate random slopes for all effects of interest [@barrRandomEffectsStructure2013]. In the semantic priming study, the items were prime-target pairs, whereas in the semantic decision and lexical decision studies, the items were individual words. In the case of interactions, random slopes were included only when the interacting variables varied within the same unit [@brauer2018a]---e.g., an interaction of two variables varying within participants (only present in Study 1). Where required due to convergence warnings, random slopes for covariates were removed, as inspired by Remedy 11 from @brauer2018a. In this regard, whereas @brauer2018a contemplate the removal of random slopes for covariates only when the covariates are not interacting with any effects of interest, we removed random slopes for covariates even if they interacted with effects of interest because these interactions were covariates themselves (covariates are indicated in the results tables).

To avoid inflating the Type I error rate (false positives), the random slopes for the effects of interest (indicated by non-shaded rows in the results tables below) were never removed [see Table 17 in @brauer2018a; for an example of this approach, see @diazNeuralSensitivityPhonological2021]. This approach arguably provides a better protection against false positives [@barrRandomEffectsStructure2013; @brauer2018a; @singmann2019a] than the practice of removing random slopes when they do not significantly improve the fit [e.g., @bernabeu2017a; @pexman2018a; @batesFittingLinearMixedeffects2015; @baayenMixedeffectsModelingCrossed2008; but also see @matuschekBalancingTypeError2017].

@@

$P$ values were calculated using the Kenward-Roger approximation for degrees of freedom [@luke2017a], in the R package 'lmerTest', Version 3.1-3 [@R-lmerTest]. The latter package in turn used 'lme4', Version 1.1-26 [@batesPackageLme42021; @batesFittingLinearMixedeffects2015]. To facilitate the convergence of the models, the maximum number of iterations was set to 1 million. Diagnostics regarding convergence and normality are provided in [\underline{Appendix B}](#appendix-B-frequentist-analysis-diagnostics). Effects that were non-significant or very small are best interpreted by considering their 95% confidence intervals [@cummingNewStatisticsWhy2014], which are shown in the results tables as well as in plots.

@@

A Bayesian analysis was performed to complement the estimates that had been obtained in the frequentist analysis. Whereas the focus of the frequentist analysis had been hypothesis testing, using $p$ values, the purpose of the Bayesian analysis was parameter estimation. Accordingly, we estimated the posterior distribution of every effect, without calculating Bayes factors [for other examples of this approach, see @milekEavesdroppingHappinessRevisited2018; @preglaVariabilitySentenceComprehension2021; @rodriguez-ferreiroSemanticPrimingSchizotypal2020; for comparisons between estimation and hypothesis testing, see @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018; @schmalzWhatBayesFactor2021; @tendeiroReviewIssuesNull2019; @tendeiroOnTheWhite2022; @rouderBayesianInferencePsychology2018; @vanravenzwaaijAdvantagesMasqueradingIssues2021]. In the estimation approach, the estimates are interpreted by considering the position of their credible intervals in relation to the expected effect size. That is, the closer an interval is to an effect size of 0, the smaller the effect of that predictor. For instance, an interval that is symmetrically centred on 0 indicates a very small effect, whereas---in comparison---an interval that does not include 0 at all indicates a far larger effect.
